---
title: "Lab 06 - Ugly charts and Simpson's paradox"
author: "George Nesbitt"
date: "02/18"
output: github_document
---

### Load packages and data

```{r load-packages, message = FALSE}
library(tidyverse) 
library(dsbox)
library(mosaicData) 
```
```{r load}
staff <- read_csv("data/instructional-staff.csv")
```


```{r intro code}
staff_long <- staff %>%
  pivot_longer(cols = -faculty_type, names_to = "year") %>%
  mutate(value = as.numeric(value))
```
```{r}
staff_long %>%
  ggplot(aes(
    x = year,
    y = value,
    group = faculty_type,
    color = faculty_type
    
  )) +
  geom_line(size=4) +
  labs(x="year", y="hires", title="Hires by type of faculty by year")
```
### Exercise 1
```{r look at data}
data(Whickham)
glimpse(Whickham)
```
I think that this data most likely came from observational studies because it would be unethical to run an experiment such as this. Especially one resulting in the death of lots of people.
### Exercise 2
There are 1,314 rows (observations) and 3 columns. Each column represents a smoker/non-smoker in whickham. 

### Exercise 3
There are three variables (columns). The first variable is outcome, this is a 


### Excercise 4
I would expect there to be a positive relationship between smoking status and death given there are serious health risk associated with the behavior.

### Excercise 5
```{r count}
Whickham %>%
  count(smoker, outcome)
```
```{r create visualization}
Whickham %>% 
  ggplot(aes(x=smoker, fill=outcome)) +
  geom_bar()
labs(
  title="Health Outcomes by smoking status",
  x="Smoking status",
  fill="Outcome"
)
```
```{r find prob}
Whickham %>% count(smoker, outcome) %>% group_by(smoker) %>% 
  mutate(probability = n /sum(n))
```

### Excercise 6
```{r create new variable}
Whickham <- Whickham %>% mutate(age_cat=case_when(
                                age <= 44 ~ "18-44",
                                age > 44 & age <= 64 ~ "45-64",
                                age > 64 ~ "65+"))
```
### Excercise 7
```{r}
Whickham %>% 
  ggplot(aes(x=smoker, fill=outcome)) +
  geom_bar() + facet_wrap(~age_cat)
labs(
  title="Health Outcomes by smoking status",
  x="Smoking status",
  fill="Outcome")
```

